CFF GCT Study

### 2025-06-25 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
GCT manual flowchart instead
GCT absolute value of days between OGTT and GCT

@@

```{r results='asis'}
#| label: table1a
#| tbl-cap: Participant characteristics by OGTT status
#| message: false
# Basic table formula
f <- ~ HbA1c + kwt(Age, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Weight (kg)`, "N", "Nmiss2", "medianq1q3", "range") + `Height (cm)` +
  kwt(BMI, "N", "Nmiss2", "medianq1q3", "range") + Gender + fe(Race) +
  fe(Ethnicity) + fe(`Pancreatic insufficient`) +
  fe(`Type 2 diabetes in the family - Father`) +
  fe(`Type 2 diabetes in the family - Mother`) +
  fe(`Brothers (number with diabetes)`) + fe(`Sisters (number with diabetes)`) +
  kwt(
    `Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)`,
    "N", "Nmiss2", "medianq1q3", "range"
  ) +
  `Highest FEV1pp` + `Second Highest FEV1pp` + `Average of Highest FEV1pp` +
  kwt(`GCT 0 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 0 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 60 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 60 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 0 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 60 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 120 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Height %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Weight %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`BMI %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`BMI Value (Adults Only)`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Days From OGTT to GCT`, "N", "Nmiss2", "medianq1q3", "range") +
  fe(Diagnosis) + fe(Dysglycemia) + fe(Diabetes) + fe(CFRD)
t1a <- tableby(update(f, `Got OGTT and GCT?` ~ . - `Got OGTT and GCT?`),
  data = df, control = mycontrols
)
summary(t1a, pfootnote = T, digits = 2)
# Limit to 189 after this
df <- df %>%
  filter(`Got OGTT and GCT?` == "Have GCT 1 Hour Plasma and OGTT 2 Hour") %>%
  select(-`Got OGTT and GCT?`)
```

@@

# Flowchart

```{r}
flow_df <- tibble(
  from = c(
    "Initial Screen with GCT",
    "Initial Screen with GCT", 
    "GCTpl >= 147",
    "GCTpl < 147"
  ),
  to = c(
    "GCTpl >= 147", 
    "GCTpl < 147", 
    "Oral Glucose Tolerance Test", 
    "Repeat Screen in 1 Year"
  )
)

ggflowchart(flow_df)
```

@@

# Nomogram

### `GCT 0 min capillary` and HbA1c

```{r}
f <- lrm(CFRD ~ `GCT 0 min capillary` + HbA1c,
  data = df
)
ddist <- datadist(df)
options(datadist = "ddist")
nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD", lp = F)
plot(nom)
```

Here’s how to use a nomogram. Pick a point on the GCT 0 min capillary and HbA1c scales and get the associated points for each. Add the points together to get the associated probability based on the Total Points line.

For example:

- GCT 0 min capillary = 130 (40 points)
- HbA1c = 6.6 (10 points)

Add up the points:

40 + 10 = 50 Total Points

50 Total Points lines up at about 0.08, so this person has approximately an 8% chance of CFRD. The process is the same for the nomograms below.

### `GCT 0 min capillary`, HbA1c, and `GCT 60 min plasma`

```{r}
f <- lrm(CFRD ~ `GCT 0 min capillary` + HbA1c + `GCT 60 min plasma`,
  data = df
)
ddist <- datadist(df)
options(datadist = "ddist")
nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD", lp = F)
plot(nom)
```

### `GCT 0 min capillary`, and `GCT 60 min plasma`

```{r}
f <- lrm(CFRD ~ `GCT 0 min capillary` + `GCT 60 min plasma`,
  data = df
)
ddist <- datadist(df)
options(datadist = "ddist")
nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD", lp = F)
plot(nom)
```

# Decision trees

## Predicting CFRD

```{r}
tree <- rpart(CFRD ~ `GCT 0 min capillary` + HbA1c + `GCT 60 min plasma`,
  data = df
)
rpart.plot(tree, type = 0)
```

We just don't have enough data to do this properly for CFRD.

## Predicting dysglycemia

```{r}
tree <- rpart(Dysglycemia ~ `GCT 0 min capillary` + `GCT 60 min plasma`, data = df)
rpart.plot(tree, type = 0)
```

>>>>

df$`Days From OGTT to GCT (Absolute Value)` <- abs(df$`Days From OGTT to GCT`)

@@

```{r results='asis'}
#| label: table1a
#| tbl-cap: Participant characteristics by OGTT status
#| message: false
# Basic table formula
f <- ~ HbA1c + kwt(Age, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Weight (kg)`, "N", "Nmiss2", "medianq1q3", "range") + `Height (cm)` +
  kwt(BMI, "N", "Nmiss2", "medianq1q3", "range") + Gender + fe(Race) +
  fe(Ethnicity) + fe(`Pancreatic insufficient`) +
  fe(`Type 2 diabetes in the family - Father`) +
  fe(`Type 2 diabetes in the family - Mother`) +
  fe(`Brothers (number with diabetes)`) + fe(`Sisters (number with diabetes)`) +
  kwt(
    `Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)`,
    "N", "Nmiss2", "medianq1q3", "range"
  ) +
  `Highest FEV1pp` + `Second Highest FEV1pp` + `Average of Highest FEV1pp` +
  kwt(`GCT 0 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 0 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 60 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 60 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 0 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 60 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 120 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Height %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Weight %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`BMI %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`BMI Value (Adults Only)`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Days From OGTT to GCT`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(
    `Days From OGTT to GCT (Absolute Value)`,
    "N", "Nmiss2", "medianq1q3", "range"
  ) +
  fe(Diagnosis) + fe(Dysglycemia) + fe(Diabetes) + fe(CFRD)
t1a <- tableby(update(f, `Got OGTT and GCT?` ~ . - `Got OGTT and GCT?`),
  data = df, control = mycontrols
)
summary(t1a, pfootnote = T, digits = 2)
# Limit to 189 after this
df <- df %>%
  filter(`Got OGTT and GCT?` == "Have GCT 1 Hour Plasma and OGTT 2 Hour") %>%
  select(-`Got OGTT and GCT?`)
```

@@

@@
